Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2016-12-31 Master Thesis/synopsis.tex
Mention subgroup considerations

Now let $A$ be free-by-finite, i.e. $A = A_\text{tor} \oplus F$ for some free abelian group $F$, and $K=\ker(f)$. Since $K\subset A$ is finite, $K$ is actually contained in $A_\text{tor}$. Therefore, we see that $A/K$ is free by finite as well, and by extension, the image $I$ of $f$. Set $m = \#B/I$ and consider the canonical map $B\stackrel{\cdot m}{\to} I \to I/I_\text{tor}$. This map has kernel $\{b\in B \mid mb \in I_\text{tor}\} = B_\text{tor}$, making $B/B_\text{tor}$ isomorphic to a subgroup of the free abelian group $I/I_\text{tor}$. Now both $I_\text{tor}$ and $B_\text{tor}/I_\text{tor} \subset B/I$ are finite, hence $B_\text{tor}$ is finite as well.
\item The exact sequence induces the exact sequence $0 \longrightarrow A'_\text{tor} \longrightarrow A_\text{tor} \longrightarrow A''_\text{tor}$. Since the surrounding groups are finite, $A_\text{tor}$ is finite as well. Using (i) we see that it suffices to show that $A/A_\text{tor}$ is free-by-finite. Consider the following commutative diagram with exact rows:
\[ \begin{tikzcd}
0 \ar[r]& A'/A'_\text{tor}\ar[r]\ar[d, "f"'] & A/A_\text{tor}\ar[r]\ar[d, "\cong"] & A/(A'+A_\text{tor}) \ar[r]\ar[d, "g"]& 0\\
0 \ar[r] & p^{-1}(A''_\text{tor})/A_\text{tor} \ar[r]& A/A_\text{tor}\ar[r] & A''/A''_\text{tor}\ar[r] & 0
\end{tikzcd} \]
Note that $f$ is injective and $g$ is surjective. Using the snake lemma, we get an isomorphism $\operatorname{coker}(f)\cong \ker(g)$. We have:
\[\ker(g) = \ker( (A/A')/(A_\text{tor}/A'_\text{tor}) \to A''/A''_\text{tor} ) = A''_\text{tor}/(A_\text{tor}/A'_\text{tor})\]
which implies that $\operatorname{coker}(f)\cong \ker(g)$ is finite. Applying (i) on $f$, we can see that $p^{-1}(A''_\text{tor})/A_\text{tor}$ is free-by-finite. Since $A''/A''_\text{tor}$ is free abelian, the short exact sequence in the second row splits, making $A/A_\text{tor}$ a direct sum of a free-by-finite group and a free group.
\item Subgroups of free-by-finite groups are free-by-finite. Now assume $A'$ is finitely generated. We consider $A'$ as a subgroup of $A$. Let $T=A_\text{tor}$. By replacing $A'$, $A$, $A''$ with $A'/(T\cap A')$, $A/T$, $A''/p(T)$ respectively, we can assume that $A$ is free and $A'$ is free of finite rank. Let $B$ be a basis of $A$. There exists a finite subset $C\subset B$ such that $A'\subset \bigoplus_{b\in C}\mathbb{Z}b$. We get an exact sequence:
\[ 0 \longrightarrow \Big(\bigoplus_{b\in C} \mathbb{Z}b\Big)\Big/A' \longrightarrow A/A' \longrightarrow \bigoplus_{b\in B\setminus C}\mathbb{Z}b \longrightarrow 0 \]
The left abelian group is finitely generated, particularly free-by-finite. The right group is free, hence also free-by-finite. By (ii), $A/A'\cong A''$ is free-by-finite as well.
\item It suffices to prove this assertion for finite groups and for $A=\mathbb{Z}$. For $A$ finite we have $nA = 0$ for $n = \#A$. Since $\mathbb{Z}$ has no nilpotent elements, the intersection of all prime ideals of $\mathbb{Z}$ is already trivial. \qedhere
\end{enumerate}
\end{proof}

>>>>

Now let $A$ be free-by-finite, i.e. $A = A_\text{tor} \oplus F$ for some free abelian group $F$, and $K=\ker(f)$. Since $K\subset A$ is finite, $K$ is actually contained in $A_\text{tor}$. Therefore, we see that $A/K$ is free by finite as well, and by extension, the image $I$ of $f$. Set $m = \#B/I$ and consider the canonical map $B\stackrel{\cdot m}{\to} I \to I/I_\text{tor}$. This map has kernel $\{b\in B \mid mb \in I_\text{tor}\} = B_\text{tor}$, making $B/B_\text{tor}$ isomorphic to a subgroup of the free abelian group $I/I_\text{tor}$. Now both $I_\text{tor}$ and $B_\text{tor}/I_\text{tor} \subset B/I$ are finite, hence $B_\text{tor}$ is finite as well.
\item The exact sequence induces the exact sequence $0 \longrightarrow A'_\text{tor} \longrightarrow A_\text{tor} \longrightarrow A''_\text{tor}$. Since the surrounding groups are finite, $A_\text{tor}$ is finite as well. Using (i) we see that it suffices to show that $A/A_\text{tor}$ is free-by-finite.

Observe the following commutative diagram with exact rows, where we consider $A'$ as a subgroup of $A$:
\[ \begin{tikzcd}
0 \ar[r]& A'/A'_\text{tor}\ar[r]\ar[d, "f"'] & A/A_\text{tor}\ar[r]\ar[d, "\cong"] & A/(A'+A_\text{tor}) \ar[r]\ar[d, "g"]& 0\\
0 \ar[r] & p^{-1}(A''_\text{tor})/A_\text{tor} \ar[r]& A/A_\text{tor}\ar[r] & A''/A''_\text{tor}\ar[r] & 0
\end{tikzcd} \]
Note that $f$ is injective and $g$ is surjective. Using the snake lemma, we get an isomorphism $\operatorname{coker}(f)\cong \ker(g)$. We have:
\[\ker(g) = \ker( (A/A')/(A_\text{tor}/A'_\text{tor}) \to A''/A''_\text{tor} ) = A''_\text{tor}/(A_\text{tor}/A'_\text{tor})\]
Since $A''_\text{tor}$ is finite, $\operatorname{coker}(f)\cong \ker(g)$ is finite, too. Applying (i) on $f$, we can see that $p^{-1}(A''_\text{tor})/A_\text{tor}$ is free-by-finite. Since $A''/A''_\text{tor}$ is free abelian, the short exact sequence in the second row splits, making $A/A_\text{tor}$ a direct sum of a free-by-finite group and a free group.
\item Subgroups of free-by-finite groups are free-by-finite. Now assume $A'$ is finitely generated. We consider $A'$ as a subgroup of $A$. Let $T=A_\text{tor}$. By replacing $A'$, $A$, $A''$ with $A'/(T\cap A')$, $A/T$, $A''/p(T)$ respectively, we can assume that $A$ is free and $A'$ is free of finite rank. Let $B$ be a basis of $A$. There exists a finite subset $C\subset B$ such that $A'\subset \bigoplus_{b\in C}\mathbb{Z}b$. We get an exact sequence:
\[ 0 \longrightarrow \Big(\bigoplus_{b\in C} \mathbb{Z}b\Big)\Big/A' \longrightarrow A/A' \longrightarrow \bigoplus_{b\in B\setminus C}\mathbb{Z}b \longrightarrow 0 \]
The left abelian group is finitely generated, particularly free-by-finite. The right group is free, hence also free-by-finite. By (ii), $A/A'\cong A''$ is free-by-finite as well.
\item It suffices to prove this assertion for finite groups and for $A=\mathbb{Z}$. For $A$ finite we have $nA = 0$ for $n = \#A$. Since $\mathbb{Z}$ has no nilpotent elements, the intersection of all prime ideals of $\mathbb{Z}$ is already trivial. \qedhere
\end{enumerate}
\end{proof}
